---
title: "pre_lab_06.Rmd"
author: "Sam Carey"
date: "2023-03-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Points to hit

1.  Review of third lab questions/problems.
2.  Demonstration of combining and merging

## Chapter 23

### Task 1: Load libraries and settings

**Task** Run the following code in the gray-colored codeblock below to load the tidyverse library and turn off scientific notation.

```{r}
# turn off sci notation
options(scipen=999)
library(tidyverse)
library(lubridate)
```

### Task 2: Load data

Let's say that we have Maryland county voter registration data from four different elections in three different files. They have a similar layout and the same number of counties (plus Baltimore City).

**Task** Load the four CSV population estimate files. Do they have the same column structure?

**Answer** They all have the same column column structure with the same headers. One thing I notice is 2018 has "UNA" and "OTH" columns switched compared to the other data frames. 

```{r}
county_voters_2016 <- read_csv("data/county_voters_2016.csv")
county_voters_2018 <- read_csv("data/county_voters_2018.csv")
county_voters_2020 <- read_csv("data/county_voters_2020.csv")
county_voters_2022 <- read_csv("data/county_voters_2022.csv")
```

### Task 3: Using bind_rows

**Task** Combine the four files into a single data frame using the bind_rows function. Add a description of what this code does to your reference notebook.

*Description: This function put all 4 data frames together into one whole data frame. 

```{r}
# bind_rows with list
county_voters_combined <- bind_rows(list(county_voters_2016, county_voters_2018, county_voters_2020, county_voters_2022))
View(county_voters_combined)
```

### Task 4: Asking a question

**Task** Run the following code to filter on Frederick County's registration data and look at the results. What stands out to you about the changes in party registration over time?

**Answer** It seems as though there was an bigger increase in registered Democrats across Frederick. Registered Republicans seem to have gradually increased, and the unaffiliated and the other columns had pretty decent increases too. 

```{r}
county_voters_combined %>% filter(COUNTY == 'Frederick')
```

## Joining data

### Task 5: Loading population data

**Task** Load the Maryland 2020 county population data 

```{r}
maryland_population <- read_csv('data/maryland_population_2020.csv')
view(maryland_population)
```

### Task 6: Join population and voter data

**Task** Run the following code to join Maryland population data and the 2022 voter registration data together using the common column county as the key. How many rows are there now? How many *should* there be? 
**Answer** There are 24 rows now, and we should have 24 rows. 

```{r}
# with nrow included to show row total
maryland_population %>% left_join(county_voters_2022, by="COUNTY") %>% nrow()
```

### Task 7: Looking at Joined Data

**Task** Examine the combined data

```{r}
# without nrow
maryland_population %>% left_join(county_voters_2022, by="COUNTY")
```

### Task 8: Assign Joined Data to New Dataframe

**Task** Create a new dataframe with population and voter data joined together. Add a description of what this code does to your reference notebook.

```{r}
maryland_population_with_voters <- maryland_population %>% left_join(county_voters_2022, by="COUNTY")
```

### Task 9: Add the Demographic Data

**Task** Load the Maryland demographic data and examine it. What kind of information does it contain?
**Answer** It contains information on race like white, Black, Asian, one race, two races, etc. It also gives numbers of the population that is 18 and over. 

```{r}
maryland_demographics <- read_csv('data/maryland_demographics.csv')
view(maryland_demographics)
```

### Task 10: Join demographic data to combined voting/population data

**Task** Join the demographic data to the our combined voting/population data. To do this, we need to specify the join columns since the names aren't the same in the two dataframes.

```{r}
maryland_population_with_voters_and_demographics <- maryland_population_with_voters %>% left_join(maryland_demographics, by=c("COUNTY"="NAME"))
```

### Task 11: Asking Demographic Questions

**Task** Using mutate, let's find the county with the highest 18+ Black population as a percentage of all population 18 and over and also the percentage of Democrats in that county. Put your answers below. 

**Answer** The county with the highest 18+ Black population as a percentage is Prince George's County with about 60.9%. The percentage of Democrats in PG County is about 77.1%.

```{r}
maryland_population_with_voters_and_demographics %>%
  mutate(pct_black_18_plus = (pop_black/pop_18_over)*100, pct_dems = (DEM/Total)*100) %>% 
  arrange(desc(pct_black_18_plus)) %>% 
  select(COUNTY, pct_black_18_plus, pct_dems)
```

### Task 12: Asking Demographic Questions

**Task** Now let's find the county with the highest 18+ Asian population as a percentage of all population 18 and over and also the percentage of Democrats in that county. How do these results compare to the previous task? Put your answers below. 

**Answer** The county with the highest 18+ Asian population as a percentage is Howard County. (Montgomery County falls right behind). There are about 52.6% Democrats in Howard County. Compared to the previous task, there is a lot less of an 18+ Asian population and Democrats in Howard County than percentage of 18+ Black population and Democrats in PG County. 

```{r}
maryland_population_with_voters_and_demographics %>%
  mutate(pct_asian_18_plus = (pop_asian/pop_18_over)*100, pct_dems = (DEM/Total)*100) %>% 
  arrange(desc(pct_asian_18_plus)) %>% 
  select(COUNTY, pct_asian_18_plus, pct_dems)
```

### Task 13: Fixing Join Problems

Sometimes joins look like they should work but don't. Often this is due to the two columns you're joining on having different data types: joining a <chr> column to a <dbl> column, for example. Let's walk through an example of that using some demographic data by zip code.

**Task** Run the following code to load the Zip Code Tabulation Area data for Maryland. What's the datatype of the ZCTA5N column? 

**Answer** The ZCTA5N column is a <dbl> column, or double, meaning a number. 

```{r}
maryland_zcta <- read_csv('data/maryland_zcta.csv')
glimpse(maryland_zcta)
```

The column we're interested in, ZCTA5N, is a <dbl> column - it's a number. We want it to be a <chr> column - text.

**Task** Run the following code to change the datatype of ZCTA5N from numeric to character. What's the datatype of ZCTA5N? 

**Answer** The datatype of ZCTA5N is now a <chr>, or character column. (Text).

```{r}
maryland_zcta <- maryland_zcta %>% mutate(ZCTA5N = as.character(ZCTA5N))
```

Now we can join this dataframe to other zip code data where the zip code column is text, not numbers.
