---
title: "Reverse Engineering Project"
author: "Samantha Carey and Dorothy Hood"
date: "2023-03-03"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Comic Books Are Still Made By Men, For Men And About Men](https://fivethirtyeight.com/features/women-in-comic-books/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

dc_comics <- read_csv("/Users/samlcarey/Desktop/j472/major_assignments/reverse_engineering/data/dc_wikia_data.csv")
  cleaned_dc_comics <- dc_comics %>%
  clean_names() 
marvel_comics <- read_csv("/Users/samlcarey/Desktop/j472/major_assignments/reverse_engineering/data/marvel_wikia_data.csv") 
  cleaned_marvel_comics <- marvel_comics %>%
  clean_names() 

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: "The characters within them are also disproportionately men, as are the new characters introduced each year."
* **Analysis summary**: It seems that in total, there are 1,967 D.C. comic female characters out of total of 6,896 D.C. comic characters, and there are 3,837 female comic Marvel characters out of 16,376 comic Marvel characters. In comparison, there are 4,783 male D.C. characters and 11,638 male Marvel characters when I switch "Female Characters" to "Male Characters" in the code below. The sentence above in the story is definitely confirmed with this data. 
```{r}
# Put code to reverse engineer sentence here
cleaned_dc_comics %>% 
  filter(sex == "Female Characters") %>% 
  group_by(year) %>% 
  arrange(desc(year)) 
cleaned_marvel_comics %>% 
  filter(sex == "Female Characters") %>% 
  group_by(year) %>% 
  arrange(desc(year))
# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here] 
* **Analysis summary**: Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-