---
title: "Reverse Engineering Project"
author: "Samantha Carey and Dorothy Hood"
date: "2023-03-03"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Comic Books Are Still Made By Men, For Men And About Men](https://fivethirtyeight.com/features/women-in-comic-books/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

dc_comics <- read_csv("/Users/samlcarey/Desktop/j472/major_assignments/reverse_engineering/data/dc_wikia_data.csv")
  cleaned_dc_comics <- dc_comics %>%
  clean_names() 
marvel_comics <- read_csv("/Users/samlcarey/Desktop/j472/major_assignments/reverse_engineering/data/marvel_wikia_data.csv") 
  cleaned_marvel_comics <- marvel_comics %>%
  clean_names() 

```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: "The characters within them are also disproportionately men, as are the new characters introduced each year."
* **Analysis summary**: It seems that in total, there are 1,967 D.C. comic female characters out of total of 6,896 D.C. comic characters, and there are 3,837 female comic Marvel characters out of 16,376 comic Marvel characters. In comparison, there are 4,783 male D.C. characters and 11,638 male Marvel characters when I switch "Female Characters" to "Male Characters" in the code below. The sentence above in the story is definitely confirmed with this data. 
```{r}
# Put code to reverse engineer sentence here
cleaned_dc_comics %>% 
  group_by(sex) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) 
cleaned_marvel_comics %>% 
  group_by(sex) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) 
# Display results of code below this codeblock
cleaned_dc_comics
cleaned_marvel_comics
```

### Sentence 2

* **Sentence text**: "Females make up about one in four comic book characters."

* **Analysis summary**: The first part of the code I searched using ChatGPT. The prompt was: "How can I use code in R to list out the total number comic book characters in both comic companies?" and "How can I combine these two companies to show all of their data together?" We then filter out the total female characters in both companies. If you take the total number of female characters, 5,804, and divide it by the total number of characters, 23,272, and multiply it by 100, you get about 24.9 or 25%, which is 1 in 4. So, this proves what's stated in the sentence. 

```{r}
# Put code to reverse engineer sentence here
total_characters <- nrow(cleaned_dc_comics) + nrow(cleaned_marvel_comics)
total_characters

dc_and_marvel <- bind_rows(cleaned_dc_comics, cleaned_marvel_comics)

dc_and_marvel %>% 
  group_by(sex) %>% 
  filter(sex == "Female Characters") %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
# Display results of code below this codeblock
dc_and_marvel

```

### Sentence 3

* **Sentence text**: "Of characters with gender data and 100 or more appearances (294 DC characters and 414 Marvel characters), only 29.0 percent of DC’s are female, and a similar 31.1 percent of the Marvel crowd is."

* **Analysis summary**: We first made new data frames to hold each character in each company and if they appeared 100 or more times. In dc_character_appearcourt, we can see 294 comic book characters, and in marvel_character_appearcourt, we can see 414. This proves the first part of the sentence. Then, we filtered out from those data frames which characters were females and appeared 100 or more times. In DC, we got 85 characters, and in Marvel we got 128. If you take 85 and divide it by 294 and multipy it by 100, you get 28.9%. If you take 128 and divide it by 414 and multiply it by 100, you get 30.9%. Although not exactly 29.0% and 31.1% like the sentence says, the numbers we got can be rounded and are extremely similar. This proves the sentence. 

```{r}
# Put code to reverse engineer sentence here
dc_character_appearcount <- cleaned_dc_comics %>%
  group_by(name, sex) %>%
  summarise(count = sum(appearances)) %>%
  filter(count >= 100) %>% 
  arrange(desc(count))
marvel_character_appearcount <- cleaned_marvel_comics %>%
  group_by(name, sex) %>%
  summarise(count = sum(appearances)) %>%
  filter(count >= 100) %>% 
  arrange(desc(count))

dc_character_appearcount %>%
  filter(sex == "Female Characters" & count >= 100) %>% 
  group_by(name) %>%
  summarise(count = sum(count)) %>%
  arrange(desc(count))
marvel_character_appearcount %>%
  filter(sex == "Female Characters" & count >= 100) %>% 
  group_by(name) %>%
  summarise(count = sum(count)) %>%
  arrange(desc(count))
# Display results of code below this codeblock
dc_character_appearcount
marvel_character_appearcount

```

### Sentence 4

* **Sentence text**: "Women and men were almost exactly equally likely to have a secret identity in the Marvel universe (49.4 percent of males and 49.5 percent of females), while in the DC universe 51 percent of males had a secret identity and 45 percent of females did."

* **Analysis summary**: After putting in this code for each comic company, it listed out the "ids," sex and count. This showed how many male and female characters in DC and Marvel had secret identities, among other genders. For Marvel, when you add up male characters that were labeled with secret identity, public identity and no dual identity, you get 9,119 male characters. When you divide the male characters with secret identities, 4,506, by 9,119 and multiply it by 100, you get 49.4%. There are a total of 3,042 females, and dividng their secret identity total (1505) by that total, you get 49.5%. 

For DC, you just have to add the secret identity and public identity totals. For males, we divided 1,751 by 3,413 and got 51%. For females, we divided 625 by 1,390 and got 45%. Overall, this code confirmed the findings in the story. 

```{r}
# Put code to reverse engineer sentence here
marvel_identities <- cleaned_marvel_comics %>%
  group_by(id, sex) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

dc_identities <- cleaned_dc_comics %>%
  group_by(id, sex) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
# Display results of code below this codeblock
dc_identities
marvel_identities
```

### Sentence 5

* **Sentence text**: "Women were substantially underrepresented among characters with at least one appearance. Among the characters for which we have gender data, females made up only 29.3 percent of the DC character list and 24.7 percent of the Marvel roster."

* **Analysis summary**: We made new data frames for each comic company, and they show DC's characters, their sex, and their number of appearances that is at least 1. For DC, there is a total of 1,880 female characters with at least one appearance. When dividing this by the total gendered characters (leaving out N/A), 6,427, and multipy it by 100, you get 29.3%. There are 3,598 total female characters with at least one appearance, and dividng that number by the total (minus N/A and "agender characters"), 14,500, and mutiply it by 100, you get 24.8% which is .1 percent off from 24.7%. Overall, this seems to confirm the stories findings.  

```{r}
# Put code to reverse engineer sentence here
dc_appearances <- cleaned_dc_comics %>%
  filter(appearances >= 1) %>%
  group_by(name, sex, appearances) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

marvel_appearances <- cleaned_marvel_comics %>%
  filter(appearances >= 1) %>%
  group_by(name, sex, appearances) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
# Display results of code below this codeblock
dc_appearances
marvel_appearances
```

-30-