---
title: "lab_03"
author: "Sam Carey"
date: "2023-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)

# Load janitor
library(janitor)

```

## Load data and clean up the column names

Load the two datasets we worked with in the pre_lab: the Maryland payments to vendors and Maryland grants and loans and, using janitor, change the columns so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows
payments <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2023.csv", guess_max=322138)
cleaned_payments <- payments %>%
  clean_names()

# Load in the Maryland grants and loan data
md_grants_loans <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv")
cleaned_md_grants_loans <- md_grants_loans %>%
  clean_names() %>% 
  rename(source = grantor)

```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed. Write code to find rows where the vendor name contains the county's school system, starting with a broad search and then narrowing it down. When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

A1: I found a total of 79 records (or about 79 records) of state agencies paying money to the PG County School system. There are a LOT of abbreviations that relate to the PG County school system. Those can be seen below with either an apostrophe added to "George's," Prince George's turned to "PG" or "PG CO," Board of Education turned to "BD of Education," etc. I counted the Board of Education of PG County because they are a part of the school system. I will note there may have been vendor names missed due to the multitude of abbreviations and different names. 

```{r}
pg_schools <- cleaned_payments %>%
  filter(vendor_name == "PRINCE GEORGES COUNTY PUBLIC SCHOOLS" | vendor_name == "BOARD OF EDUCATION OF PRINCE GEORGES COUNTY MD" | vendor_name == "PRINCE GEORGES PUBLIC SCHOOLS" | vendor_name == "PRINCE GEORGES SCHOOL" | vendor_name == "BOARD OF EDUCATION OF PRINCE GEORGE'S COUNTY" | vendor_name == "PG COUNTY PUBLIC SCHOOL" | vendor_name == "PRINCE GEORGES COUNTY SCHOOL" | vendor_name == "BD OF EDUCATION OF PG CO")
```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? What do you notice about the years listed?

A2: The fiscal year 2022 had the largest total amount. I noticed that all the fiscal years listed are in the 2000s, from 2008 to 2023. 2023 has the lowest total amount (so far). 

```{r}
pg_schools %>%
  group_by(fiscal_year) %>% 
  summarize(total_amount = sum(amount)) %>% 
  arrange(desc(total_amount))
```
Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received. Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

A3. One vendor that has similar names is "AYERS SAINT GROSS INC," which is spelled like that and also as "AYERS SAINT GROSS" and "AYERS SAINT GROSS INCORPORATED." Another vendor name is "BECTON DICKINSON" which is also called "BECTON DICKINSON AND COMPANY." A third vendor name is "A R MARANI INC" which is also called "A.R. MARANI, INC." If I was to clean the data, I'd possibly use the function get_dupes(). This could show all vendors that possibly have similar or basically the same names, and then use distinct() to get rid of the duplicate rows. Also, maybe there's a way to take a key word from the vendor names, like "AYERS" for example, and have all of those vendor name variations fall under that one key word to better organize everything, and do that for all of the vendors that have similar names listed.  

```{r}
cleaned_payments %>%
  filter(str_detect(agency_name, "CAPITAL PROJECTS- UNIVERSITY OF MARYLAND"))
```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which columns contain the information you want, your code should standardize the data needed to find just the grants for stem cell research in a single filter. Then create a total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

A4: In the year 2009, there was the most amount of grant money put towards stem cell research projects. It seems as the years went on the grant money became less, with 2021 having the least amount of grant money towards stem cell research projects. 

```{r}
stem_cell_research_grants <- cleaned_md_grants_loans %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, "stem cell research grant")) %>% 
  group_by(fiscal_year) %>% 
  summarize(total_amount = sum(amount)) %>% 
  arrange(desc(total_amount))
  
```
