---
title: "lab_11"
author: "Sam Carey"
date: "2023-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* Our usual libraries for working with data, plus rvest.

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
library(rvest)
library(tidyverse)
library(janitor)
library(lubridate)
library(stringr)
```

Let's get to scraping.

## Questions

**Q1**. Scrape the listing of available Maryland state grants at https://grants.maryland.gov/Pages/StateGrants.aspx into a dataframe. You should have three columns, one of which is a date, so make sure the date column has a date datatype. Then, write code to count the number of grants opportunities offered by each organization listed in your dataframe, showing the organization with the most grant opportunities first. Which state agency has the most?

**A1** According to the data frame, the state agency that has the most grant opportunities is the Maryland Energy Administration with 14. The data is interesting, and one thing to note is that there aren't dates for every single grant.

A big piece to note for this question, however, is 14 may not be the total number of grants offered by the Maryland Energy Administration, along with a few other agencies. There are abbreviated names for agencies that are the same. The Maryland Energy Administration is also called "MD Energy Administration" and "Energy Administration." Realistically, there should be about 18 grants for this agency. These should all be together. You can also see this with the Maryland State Arts Council, which is also, in the data, called "MD Arts Council" and "MD State Arts Council."

```{r}
grants_url <- "https://grants.maryland.gov/Pages/StateGrants.aspx"

results <- grants_url %>%
  read_html() %>% 
  html_table()

results <- results[[1]]

results <- results %>%
  clean_names() %>%
  mutate(due_date = mdy(due_date))

state_grants <- results %>% 
  group_by(organization) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```

**Q2** Next, let's scrape the list of press releases from Maryland's Office of the Public Defender, https://www.opd.state.md.us/press-releases. This isn't a table, so you'll need to use `html_elements()` and your browser's inspector and do some clean up on the results. The result should be a dataframe with two columns that contain the date and title, and the date column should have a date datatype. The challenge here is figuring out how to isolate the releases.

When you finish scraping into a dataframe, write code to find the press releases that have the word "police" in the title. How many are there and when was the most recent one?

**A2** There are 9 press releases that have the word "Police" in the title. The most recent one in June 21, 2021. (2021-06-21)

```{r}
opd_url <- "https://www.opd.state.md.us/press-releases"

opd_results <- opd_url %>%
  read_html()

opd_results %>% html_elements('a')

opd_results %>% html_elements('a') %>% html_text()

releases <- opd_results %>% html_elements('a') %>% html_text() %>% as_tibble()

releases <- opd_results %>% html_elements('a.wixui-rich-text__text') %>% html_text() %>% as_tibble()

releases <- opd_results %>% html_elements('a.wixui-rich-text__text') %>% html_text() %>% as_tibble() %>% separate(value, c('date', 'title'), sep=":")

releases <- opd_results %>% html_elements('a.wixui-rich-text__text') %>% html_text() %>% as_tibble() %>% separate(value, c('date', 'title'), sep=":") %>% drop_na(title) %>% mutate(title = str_squish(title), date = mdy(date))

releases_police <- releases %>% 
  filter(str_detect(title, "Police"))
  
```

**Q3** Sen. Ben Cardin, D-Maryland, has posted hundreds of press releases at https://www.cardin.senate.gov/?post_type=press-releases. It would be great to have all of them in a dataframe that has the following columns: date, title and url.

To do this, you will need to scrape the page's html and save that to a variable, and _then_ extract the dates, titles and urls into _separate_ dataframes using html_elements(). And remember how we turn a list into a dataframe. The function `html_text()` pulls out the contents of a tag, but for urls we want the HTML attribute. Rvest gives you a way to extract the URL from a link; google to find out what it is.

At the end, you'll have three dataframes that you want to combine into a single dataframe. When we want to combine the rows of identical dataframes, we used `bind_rows()`. If you were combining columns instead of rows, there's a similar function. Find out what it is and use it to put all of the dataframes together into a single one.

When you're done, rename the columns so they make sense, then make sure the date column is an actual date.

Finally, tell me what questions you could ask of this data. Be creative.

**A3** 1. What are these press releases about? What does Senator Cardin find important to release?
2. What was the differences between press release content from after the 2020 midterm elections compared to after the 2022 midterm elections? 
3. Did he publish more promises or announcements of funding or other topics to gain voters before each election?
4. Did the amount of press releases he published and continues to publish keep contact with his constituents and improve his popularity as a state Senator? Did it get him more votes?
5. What kind of policy did he announce when COVID-19 was at it peak?

```{r}
sen_url <- "https://www.cardin.senate.gov/?post_type=press-releases"

# Dates first
sen_results <- sen_url %>%
  read_html()

sen_results %>% html_elements('h5')

sen_results %>% html_elements('h5') %>% html_text()

release_dates <- sen_results %>% html_elements('h5') %>% html_text() %>% as_tibble()

release_dates <- release_dates %>% 
  rename(date = value) %>% 
  mutate(date = mdy(date)) 

# Titles second
sen_results <- sen_url %>%
  read_html()

sen_results %>% html_elements('a')

sen_results %>% html_elements('a') %>% html_text()

release_titles <- sen_results %>% html_elements('a') %>% html_text() %>% as_tibble()

release_titles <- sen_results %>% html_elements('h3.customBlog_item__title') %>% html_text() %>% as_tibble() %>% rename(title = value) %>% mutate(title = str_squish(title))

# URLs

sen_results <- sen_url %>%
  read_html()

release_links <- sen_results %>% 
  html_elements("h3.customBlog_item__title a") %>% 
  html_attr("href") %>% 
  as_tibble() %>% 
  rename(url = value)

# Combine
combined_releases <- cbind(release_dates, release_titles, release_links)

```
