---
title: "lab_10"
author: "Sam Carey"
date: "2023-04-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* Our usual libraries, plus some of the map-specific ones from the pre_lab. You'll also be fetching Census data in Q3. Read the questions carefully!

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Load libraries
library(tidyverse)
library(sf)
library(janitor)
library(tidycensus)
library(tigris)
#census_api_key("549950d36c22ff16455fe196bbbd01d63cfbe6cf")
```

Let's explore Maryland elections, power outages and electric vehicle data and make some maps! 

## Questions 

**Q1.** Maryland voters made history in November by electing the state's first two African-American statewide office holders: Wes Moore as governor and Anthony Brown as attorney general. You want to compare the two's electoral performances. Using the `md_statewide_county.csv` file in the data folder, calculate the total number of votes for both the governor's race and the attorney general's race, then percentages for both Moore and Brown _based on the total of their respective contests_. Then calculate the difference between their percentages.

Get the geometry for Maryland's counties using tigris and, joining that dataframe to your results dataframe, make a county-level map of the differences between Moore and Brown's percentages. What is the story here, and what is the county with the most interesting results?

**A1.** First, it seems here that PG County and Baltimore City both show the smallest differences in vote percentages for Brown and Moore. Then, it shows the counties with the largest percent differences found in Carroll County, Kent County and Queen Anne's County. These three counties I think have the most interesting results because, why was there such a difference between voting performances between Moore and Brown in these counties? Checking the data, it shows Moore performed better than Brown. I think the story could be here, why is that the case? Why did Moore perform better than Brown in these counties?

```{r}
md_statewide_county <- read_csv("data/md_statewide_county.csv")

md_statewide_county_votes <- md_statewide_county %>% 
  mutate(total_votes_gov = Cox + Moore + Lashar + Wallace + Harding + Sefcik + GovWritein) %>% 
  mutate(total_votes_ag = Peroutka + Brown + AgWritein) %>% 
  mutate(pctmoore = Moore/total_votes_gov *100) %>% 
  mutate(pctbrown = Brown/total_votes_ag *100) %>% 
  mutate(pct_differences = pctmoore - pctbrown)

counties <- counties()

md_counties <- counties %>%
  filter(STATEFP == "24")

md_statewide_county_votes <- md_statewide_county_votes %>% 
  mutate(GEOID = as.character(GEOID))

md_counties_and_votes_geom <- md_counties %>% 
  inner_join(md_statewide_county_votes, by = "GEOID")

ggplot() +
  geom_sf(data=md_counties) +
  geom_sf(data=md_counties_and_votes_geom, aes(fill=pct_differences)) +
  scale_colour_viridis_b(option="magma") +
  theme_minimal()

```

**Q2.** Make a county-level map to explore the number of customers impacted by power outages in Maryland. Load the `county_outages_march2023.csv` file in the data folder and calculate a rate of customers with outages per 10,000 total customers for each county. Then join that dataframe to your Maryland counties dataframe you made in Q1 and create a county-level map of the results; you may need to tweak the fill scale to make the distinctions clearer. What's the potential story here? Given access to other months' outage data, what would your next steps be in reporting a story?

**A2.** First, it seems that Washington County and Frederick County experienced the least amount of power outages per 10,000. It then is obviously shown that Talbot County experienced the most outages per 10,000. The potential story could be what is causing the large amount of power outages Talbot County is facing, and/or what are Washington and Frederick County doing that they have experienced little power outages? If I had access to other months' outage data, I would study the trend over time how these counties experienced more or less outages, and keep track of the counties with the most and counties with the least. Then we could try to find out why certain counties are dealing with this more. 

```{r}
county_outages_march2023 <- read_csv("data/county_outages_march2023.csv")

march23_outages_rates <- county_outages_march2023 %>% 
  mutate(outages_per_10000 = (outage_customers / total_customers) * 10000)

march23_outages_and_counties <- md_counties %>% 
   inner_join(march23_outages_rates, by = c('NAME' = 'county'))

ggplot() +
  geom_sf(data=md_counties) +
  geom_sf(data=march23_outages_and_counties, aes(fill=outages_per_10000)) +
  scale_fill_viridis_b(option="magma",trans = "log") +
  theme_minimal()
  
```

**Q3.** You've been assigned a story about the use of electric vehicles in Maryland and you want to make a map of the prevalence of electric vehicles in each county. Load the `ev_2020_2023.csv` file, which has the number of electric and hybrid vehicles per county from July 2020 and March 2023 and calculate a percentage change for each county (remember, that's `new-old/old`). Using the Maryland counties dataframe you made for Q1, make a county map showing the percentage change.What counties stand out to you? What questions do you have about the changes you're seeing?

Next, use tidycensus to retrieve county population data for Maryland from 2021 (you did this in the pre_lab) and join it to your vehicles dataframe. HINT: you may not need to join it on the county name, and you should already have the geometry!

Then, calculate a per capita rate of electric & hybrid vehicles using the March 2023 figure and the population estimate. You will need to choose the "per capita" value.

Finally, make a second county map showing the per capita rate. What counties stand out on this map? How is it different from the first one you made?

**A3.** From the first map mapping percentage change for each county, it seems Dorchester County has the lowest percent change of electric and hybrid vehicles, while Caroline County has the highest percent change. Some questions could be:

Why did Dorchester only have a small increase in their electric/hybrid vehicle usage?

Why does Caroline have a large increase in their electric/hybrid vehicle usage?

What's going on in these two areas that enabled more drivers to use these vehicles (or less drivers in Dorchester's case)?

In the second map, the county with the least per capita rate (per 10,000 people) of electric/hybrid vehicles seems to be Somerset County. The counties with the highest per capita rate (per 10,000 people) seem to be Montgomery County and Howard County. 

This map is different from my first map because it's showing completely different counties sticking out than the first map. (Dorchester and Caroline versus Somerset, Montgomery and Howard). The per capita rates also do not hit as high of numbers as the percent change numbers do (236.9 versus 425.9). This goes to show the difference in calculations, as percent change is, in this case, an increase or decrease of electric/hybrid vehicle usage over two periods of time (July 2020 and March 2023), whereas pert capita rate is, in this case, how many electric/hybrid vehicles are in use per 10,000 people in each county. It's interesting to see how these two calculations lead different results within the counties. 

```{r}
# Part one of question
ev_2020_2023 <- read_csv("data/ev_2020_2023.csv")

ev_20_23_pctchange <- ev_2020_2023 %>% 
  mutate(pctchange = (march_2023 - july_2020) / july_2020 * 100)

ev_and_counties_geom <- md_counties %>% 
   inner_join(ev_20_23_pctchange, by = c('NAME' = 'county'))

ggplot() +
  geom_sf(data=md_counties) +
  geom_sf(data=ev_and_counties_geom, aes(fill=pctchange)) +
  scale_fill_viridis_b(option="magma",trans = "log") +
  theme_minimal()

# Part two

md_county_population <- get_acs(geography = "county",
              variables = c(population = "B01001_001"),
              year = 2021,
              state = "MD")

ev_and_populations <- ev_and_counties_geom %>% 
   inner_join(md_county_population, by = "GEOID")

ev_and_populations <- ev_and_populations %>% 
  mutate(per_capita = (march_2023 / estimate) * 10000)

ggplot() +
  geom_sf(data=md_counties) +
  geom_sf(data=ev_and_populations, aes(fill=per_capita)) +
  scale_fill_viridis_b(option="magma",trans = "log") +
  theme_minimal()

```

